### 2024-09-02 cognitive-bias.qmd
20240902 - anchoring and adjustment

The *anchoring and adjustment heuristic* refers to the tendency for a person's judgments or predictions to be made with a reference point—an anchor—as a starting point from which they adjust their estimates upward or downward.
The anchor is often inaccurate and given too much weight in the person's calculation, and too little adjustment is made to the anchor.
For instance, a manager is trying to predict how many fantasy points a top Running Back may score.
The player scored 300 fantasy points last season, but the team added a stronger backup Running Back and changed the Offensive Coordinator to be a more pass-heavy offense.
The manager may use 300 fantasy points as an anchor (based on the player's performance last season), and may adjust downward 15 points to account for the offseason changes.
However, it is possible that this downward adjustment is insufficient to account not only for the offseasons changes but also for potential [regression effects](#sec-fallaciesRegression).
[Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.

>>>>

The *anchoring and adjustment heuristic* refers to the tendency for a person's judgments or predictions to be made with a reference point—an anchor—as a starting point from which they adjust their estimates upward or downward.
The anchor provides a suggestion to the person—even without their paying attention, leading to a priming effect which evokes information that is compatible with the suggestion (i.e., a suggestive influence).
We are more suggestible than we would like to think we are.
The anchor is often inaccurate and given too much weight in the person's calculation, and too little adjustment is made away from the anchor.
As an example, insufficient adjustment from an anchor may lead someone to drive too fast after exiting a highway [@Kahneman2011].
The anchoring effect occurs even when the anchor (i.e., suggestion) is a random number and is completely uninformative of the correct answer [@Kahneman2011]!

Applied to fantasy football, a manager is trying to predict how many fantasy points a top Running Back may score.
The player scored 300 fantasy points last season, but the team added a stronger backup Running Back and changed the Offensive Coordinator to be a more pass-heavy offense.
The manager may use 300 fantasy points as an anchor (based on the player's performance last season), and may adjust downward 15 points to account for the offseason changes.
However, it is possible that this downward adjustment is insufficient to account not only for the offseasons changes but also for potential [regression effects](#sec-fallaciesRegression).
[Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.

To counter the effect of the anchoring heuristic, @Kahneman2011 suggests deliberately thinking the opposite of the anchor.
And, if a person you are negotiating with makes an outrageous initial offer, instead of making an equally outrageous counter-offer, @Kahneman2011 suggests storming out or threatening to leave, and indicating that you are unwilling to negotiate with that number on the table.
This technique could be helpful during trade negotiations for players in fantasy football.
Other techniques include, instead of focusing on the anchor, focusing on the minimal offer that the opponent would accept, or on the costs to the opponent of failing to reach an agreement [@Kahneman2011].
